---
title: "Untitled"
author: "DEE"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Sampling Census Data

Because census data is organzied with person-level records within a household unit, 
we cannot just randomly sample a person-level file.



The process is not terribly complex, but may be new to some users. The steps are:

1. Create a sample of household IDs
2. Filter the full dataset down to only those hosueholds.


Below we outline the approach both with baseR as well as using the dplyr package.
Results should be similar but not identical as they are ultimately using separate
sampling functions.



Be sure to `set.seed()` to ensure the random sampling is repeatable in the future.


## Load dataset

See the code below to run this on the included `ex_census_data`, or on a true
IPUMS data extract.

```{r}
library(ipumsED)
orig_data <- ex_census_data

```


```{r, eval = FALSE}

library(ipumsr)

path_to_data <- file.path()
path_to_ddi <- file.path()


ddi <- ipumsr::read_ipums_ddi(path_to_ddi)
orig_data <- ipumsr::read_ipums_micro(ddi,data_file = path_to_data)

```



## Random sample of households - base R


Because the base R `sample()` function requires an integer size to sample, we must
first calculate the `n_size`. We do this by finding the `round`ed `length` of `unique_hhs`.

Notice that we create `unique_hhs_int`, an integer vector of household. This is
makes it easier to subset `unique_hhs`.



```{r}

## find unique hhs
unique_hhs <- unique(orig_data$SERIAL)
## create an integer vector of unique hhs
unique_hhs_int <- seq_along(unique_hhs)

## find the number of hhs to ultimately sample
n_size <- round(length(unique_hhs)*.01)



set.seed(13)
sampled_hhs_1 <- unique_hhs[sample(unique_hhs_int, size = n_size)]
sampled_dat_1 <- orig_data[orig_data$SERIAL %in% sampled_hhs_1,]


```


## Random sammle of households - dplyr

```{r}


library(dplyr)

## note: seed must be passed in the SAME CALL it is needed in.
set.seed(13)
sampled_hhs_2 <- orig_data %>% distinct(SERIAL) %>% slice_sample(prop = .01) %>% pull(SERIAL)
sampled_dat_2 <- orig_data %>% filter(SERIAL %in% sampled_hhs_2)

```



## Compare results

```{r}

table(sampled_hhs_1==sampled_hhs_2)
identical(sampled_hhs_1,sampled_hhs_2)

class(sampled_hhs_1)
class(sampled_hhs_2)

```


## Write out results

```{r, eval = FALSE}

write.csv(sampled_dat_1, "sampled_census_data.csv",row.names = FALSE)

```
